Exploratory Data Analysis With R by David Eraso
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(gridExtra)
library(GGally)
library(dplyr)
library(lattice)
library(memisc)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
red_wines <- read.csv("C:/Users/David/Documents/GitHub/Data-Analysis-P-R_EDA/wineQualityReds.csv")
```
Red Wine Quality Dataset:
Reference: P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
  Modeling wine preferences by data mining from physicochemical properties.
  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
                [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf
                [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib

# Univariate Plots Section
Number of variables (columns - tidy data)
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_ncol}
ncol(red_wines)
```

Number of observations (rows - tidy data)
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_nrow}
nrow(red_wines)
```

Variables in the dataset:
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_names}
names(red_wines)
```

Structure of the dataset:
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_str}
str(red_wines)
```

First 4 rows of the dataset:
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_head}
head(red_wines, n = 4L)
```

Summary of the dataset:
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_summary}
summary(red_wines)
```

The most important variable is 'quality'. I will begin exploring this variable with a histogram. This variable is type 'integer' and it is a score between 0 and 10, so the x axis is scaled accordingly:
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_quality_histogram}
ggplot(aes(x = quality), data = red_wines) +
  geom_histogram(colour=I("#330033"), fill=I("#990000")) +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 1)) +
  ggtitle("Quality Histogram")
  
```

Most wines received a score of 5 or 6 (over 600 each), The third most common score is 7, and scores 0, 1, 2, 9, and 10 don't seem to get any count. To verify the actual frequency I will table this variable:
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_quality_table}
table(red_wines$quality)
```

Setting the binwidth, and only taking into account scores with count greater than 0:
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_quality_histogram_bin}
ggplot(aes(x = quality), data = red_wines) +
  geom_histogram(binwidth = 0.1, 
                 colour=I("#330033"), fill=I("#990000")) +
  scale_x_continuous(limits = c(3, 9), breaks = seq(3, 8, 1)) +
  ggtitle("Quality Histogram")
  
```

Now I will explore the distributions of the other variables. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_fixed.acidity_1}
ggplot(aes(x = fixed.acidity), data = red_wines) +
  geom_histogram(colour="black", fill=I("#FF3300")) +
  ggtitle("Fixed Acidity Histogram 1")

```

Adjusting binwidth
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_fixed.acidity_2}
ggplot(aes(x = fixed.acidity), data = red_wines) +
  geom_histogram(binwidth = 0.1, colour="black", fill=I("#FF3300")) +
  ggtitle("Fixed Acidity Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_volatile.acidity_1}
ggplot(aes(x = volatile.acidity), data = red_wines) +
  geom_histogram(colour="black", fill=I("#FF9933")) +
  ggtitle("Volatile Acidity Histogram 1")

```

Adjusting binwidth
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_volatile.acidity_2}
ggplot(aes(x = volatile.acidity), data = red_wines) +
  geom_histogram(binwidth = 0.01, colour="black", fill=I("#FF9933")) +
  ggtitle("Volatile Acidity Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_citric.acid_1}
ggplot(aes(x = citric.acid), data = red_wines) +
  geom_histogram(colour="black", fill=I("#66FF33")) +
  ggtitle("Citric Acid Histogram 1")

```

Adjusting binwidth and omitting top 4% of citric acid values
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_citric.acid_2}
ggplot(aes(x = citric.acid), data = subset(red_wines, citric.acid <= quantile(citric.acid, probs = 0.96))) +
  geom_histogram(binwidth = 0.01, colour="black", fill=I("#66FF33")) +
  ggtitle("Citric Acid Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_residual.sugar_1}
ggplot(aes(x = residual.sugar), data = red_wines) +
  geom_histogram(colour="black", fill=I("#FFFF99")) +
  ggtitle("Residual Sugar Histogram 1")

```

Adjusting binwidth and omitting top 4% of residual sugar values
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_residual.sugar_2}
ggplot(aes(x = residual.sugar), data = subset(red_wines, residual.sugar <= quantile(residual.sugar, probs = 0.96))) +
  geom_histogram(binwidth = 0.1, colour="black", fill=I("#FFFF99")) +
  ggtitle("Residual Sugar Histogram 2")

```

log transform
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_residual.sugar_3}
ggplot(aes(x = log10(residual.sugar + 1)), data = subset(red_wines, residual.sugar <= quantile(residual.sugar, probs = 0.96))) +
  geom_histogram(binwidth = 0.01, colour="black", fill=I("#FFFF99")) +
  ggtitle("Log Transformed Residual Sugar Histogram 3")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_chlorides_1}
ggplot(aes(x = chlorides), data = red_wines) +
  geom_histogram(colour="black", fill=I("#99CCFF")) +
  ggtitle("Chlorides Histogram 1")

```

Adjusting binwidth and omitting top 4% of chlorides values
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_chlorides_2}
ggplot(aes(x = chlorides), data = subset(red_wines, chlorides <= quantile(chlorides, probs = 0.96)) ) +
  geom_histogram(binwidth = 0.002, colour="black", fill=I("#99CCFF")) +
  ggtitle("Chlorides Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_free.sulfur.dioxide_1}
ggplot(aes(x = free.sulfur.dioxide), data = red_wines) +
  geom_histogram(colour="black", fill=I("#85ADAD")) +
  ggtitle("Free Sulfur Dioxide Histogram 1")

```

Adjusting binwidth
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_free.sulfur.dioxide_2}
ggplot(aes(x = free.sulfur.dioxide), data = red_wines) +
  geom_histogram(binwidth = 1, colour="black", fill=I("#85ADAD")) +
  ggtitle("Free Sulfur Dioxide Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_total.sulfur.dioxide_1}
ggplot(aes(x = total.sulfur.dioxide), data = red_wines) +
  geom_histogram(colour="black", fill=I("#527A7A")) +
  ggtitle("Total Sulfur Dioxide Histogram 1")

```

Adjusting binwidth and omitting top 4% of total sulfur dioxide values
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_total.sulfur.dioxide_2}
ggplot(aes(x = total.sulfur.dioxide), data = subset(red_wines, total.sulfur.dioxide <= quantile(total.sulfur.dioxide, probs = 0.96))) +
  geom_histogram(binwidth = 1, colour="black", fill=I("#527A7A")) +
  ggtitle("Total Sulfur Dioxide Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_density_1}
ggplot(aes(x = density), data = red_wines) +
  geom_histogram(colour="black", fill=I("#99CCFF")) +
  ggtitle("Density Histogram 1")

```

Adjusting binwidth
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_density_2}
ggplot(aes(x = density), data = red_wines) +
  geom_histogram(binwidth = 0.0001, colour="black", fill=I("#99CCFF")) +
  ggtitle("Density Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_pH_1}
ggplot(aes(x = pH), data = red_wines) +
  geom_histogram(colour="black", fill=I("#FFFFCC")) +
  ggtitle("pH Histogram 1")

```

Adjusting binwidth
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_pH_2}
ggplot(aes(x = pH), data = red_wines) +
  geom_histogram(binwidth = 0.01, colour="black", fill=I("#FFFFCC")) +
  ggtitle("pH Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_sulphates_1}
ggplot(aes(x = sulphates), data = red_wines) +
  geom_histogram(colour="black", fill=I("#993399")) +
  ggtitle("Sulphates Histogram 1")

```

Adjusting binwidth and omitting top 4% of sulphates values
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_sulphates_2}
ggplot(aes(x = sulphates), data = subset(red_wines, sulphates <= quantile(sulphates, probs = 0.96))) +
  geom_histogram(binwidth = 0.01, colour="black", fill=I("#993399")) +
  ggtitle("Sulphates Histogram 2")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_alcohol_1}
ggplot(aes(x = alcohol), data = red_wines) +
  geom_histogram(colour="black", fill=I("#CCFFFF")) +
  ggtitle("Alcohol Histogram 1")

```

Adjusting binwidth
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_alcohol_2}
ggplot(aes(x = alcohol), data = red_wines) +
  geom_histogram(binwidth = 0.1, colour="black", fill=I("#CCFFFF")) +
  ggtitle("Alcohol Histogram 2")

```

log transform
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_alcohol_3}
ggplot(aes(x = log10(alcohol + 1)), data = red_wines) +
  geom_histogram(binwidth = 0.005, colour="black", fill=I("#CCFFFF")) +
  ggtitle("Log Tranformed Alcohol Histogram 3")

```




# Univariate Analysis

### What is the structure of your dataset?
There are 11 numeric variables that describe different characteristics of the 1599 red wines in the data set. 'quality' is a categorical variable that represents the score of a given red wine according to an expert.

### What is/are the main feature(s) of interest in your dataset?
Quality scores have very little variance; most wines in the dataset have a score of 5 or 6 (1319 out of 1599 - 82.5%). This is the only qualitative variable. 

Most variables have a somewhat normal distribution, usually with longer right tails, with the exception of ctric acid which has a large count for value 0, and a volatile distribution. It is also worth noting the 'well behaved' distributions for both density and ph.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
The idea is to explore how the 11 quantitative variables relate to 'quality'. Also it will be interesting to see how these variables interact with each other, especially the group of fixed acidity, volatile acidity and citric acid (ph also may be related to this group). And the group of free sulfur oxide, total sulfur oxide, and maybe sulphates.  

### Did you create any new variables from existing variables in the dataset?
No.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
The first long right tail was for residual sugar. The difference between the log transform and a histogram that omits the top 4% of the values, aside from the various gaps in the log transform, does not give additional insights. 

The histogram for alcohol also was log transformed, but again, aside shortening the tail in a similar way than omitting the top 4%, it does not give any further insights.


# Bivariate Plots Section

Corralations table for the dataset
```{r echo=FALSE, message=FALSE, warning=FALSE, Correlations}
cor(red_wines)
```
...

Pairs using initial intuition, I will look into pairwise relationship for fixed acidity, volatile acidity, citric acid, ph, and quality.
```{r echo=FALSE, message=FALSE, warning=FALSE, ggpairs_1}
ggpairs(red_wines, columns = c(2,3,4,10,13))
```

This plot is not very helpful. I will adjust the previous plot so that 'quality' is in the x axis, I will add an alpha parameter to deal with overplotting, and a smooth parameter to the scatterplots. 
```{r echo=FALSE, message=FALSE, warning=FALSE, ggpairs_adjusted}
ggpairs(red_wines, columns = c(2,3,4,10,13), 
        upper = list(continuous = "smooth", 
                     params = c(alpha = 0.1)),
        lower = list(continuous = "cor"))
```

The following plot looks for pairwise relationships between free sulfur dioxide, total sulfur dioxide, sulphates and quality.
```{r echo=FALSE, message=FALSE, warning=FALSE, ggpairs_2}
ggpairs(red_wines, columns = c(7,8,11,13), 
        upper = list(continuous = "smooth", 
                     params = c(alpha = 0.1)),
        lower = list(continuous = "cor"))
```

The following plot looks for pairwise relationships between residual sugar, chlorides, density, alcohol and quality.
```{r echo=FALSE, message=FALSE, warning=FALSE, ggpairs_3}
ggpairs(red_wines, columns = c(5,6,9,12,13), 
        upper = list(continuous = "smooth", 
                     params = c(alpha = 0.1)),
        lower = list(continuous = "cor"))
```

Finally, taking into account the correlation table for the dataset, and that the main variable for this analysis is 'quality' I will look for pairwise relationships between quality, alcohol, volatile acidity, sulphates, and total sulfur dioxide. Citric acid, although relevant looking at its correlation with quality is not considered since, anticipating the construction of a model to predict quality, I will be looking for independent variables, and citric acid is correlated to volatile acidity (which is already included).
```{r echo=FALSE, message=FALSE, warning=FALSE, ggpairs_4}
ggpairs(red_wines, columns = c(3,8,11,12,13), 
        upper = list(continuous = "smooth", 
                     params = c(alpha = 0.1,colour="blue")),
        lower = list(continuous = "cor"))
```

Now, I will look closer into distributions of variables of interest for different qualities.

First Alcohol:
A scatterplot using jitter, and the alpha parameter:
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_1}
ggplot(data = red_wines, aes(x = factor(quality), y = alcohol)) +
  geom_jitter(aes(colour = factor(quality), alpha = 0.1)) +
  ggtitle("Alcohol by Quality Scatterplot")
```

Histograms of alcohol for different qualities:
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_2}
ggplot(data = red_wines, aes(x = alcohol, fill = factor(quality))) +
  geom_bar(binwidth = 0.1) +
  ggtitle("Alcohol Histogram by Quality")
```
There are scores for wide values of alcohol, since most quality scores are 5 and 6, these two histograms take most of the plot area and there is not a clear cut insight. I will try boxplots. 

For the boxplots by quality I turn 'quality', which has 'int' type, to a categorical varible (Ref: http://stackoverflow.com/questions/16639484/how-to-convert-integer-into-categorical-data-in-r). Also for coloring the boxplot (Ref: http://www3.nd.edu/~steve/computing_with_data/11_geom_examples/ggplot_examples.html)
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_3}
ggplot(data = red_wines, aes(x = factor(quality), y = alcohol)) +
  geom_boxplot(aes(fill = factor(quality))) +
  ggtitle("Alcohol Boxplots by Quality")
```
This plot is more insightful. 

I will try now conditional medians. For the plot (Ref: http://www.cookbook-r.com/Graphs/Shapes_and_line_types/)
```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_medians_1}
alcohol.by_quality <- red_wines %>%
  group_by(quality) %>%
  summarise(alcohol_median = median(alcohol),
            alcohol_mean = mean(alcohol),
            n = n())
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_means_plot_1}
ggplot(data = alcohol.by_quality, aes(x = quality, y = alcohol_median)) +
  geom_line(linetype = 5, colour = "red", size = 2) +
  ggtitle("Alcohol conditional means by Quality")
```

Despite the decrease from score 4 to score 5, it seems that in general, red wines with higher alcohol get higher scores.

Now, other variables of interest:

Volatile Acidity:
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_4}
ggplot(data = red_wines, aes(x = factor(quality), y = volatile.acidity)) +
  geom_boxplot(aes(fill = factor(quality))) +
  ggtitle("Volatile Acidity Boxplots by Quality")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_medians_2}
volatile.acidity.by_quality <- red_wines %>%
  group_by(quality) %>%
  summarise(volatile.acidity_median = median(volatile.acidity),
            volatile.acidity_mean = mean(volatile.acidity),
            n = n())
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_means_plot_2}
ggplot(data = volatile.acidity.by_quality, aes(x = quality, y = volatile.acidity_median)) +
  geom_line(linetype = 5, colour = "orange", size = 2) +
  ggtitle("Volatile Acidity conditional means by Quality")
```

Sulphates:
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_5}
ggplot(data = red_wines, aes(x = factor(quality), y = sulphates)) +
  geom_boxplot(aes(fill = factor(quality))) +
  ggtitle("Sulphates Boxplots by Quality")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_medians_3}
sulphates.by_quality <- red_wines %>%
  group_by(quality) %>%
  summarise(sulphates_median = median(sulphates),
            sulphates_mean = mean(sulphates),
            n = n())
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_means_plot_3}
ggplot(data = sulphates.by_quality, aes(x = quality, y = sulphates_median)) +
  geom_line(linetype = 5, colour = "blue", size = 2) +
  ggtitle("Sulphates conditional means by Quality")
```

Total Sulfur Dioxide:
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_6}
ggplot(data = red_wines, aes(x = factor(quality), y = total.sulfur.dioxide)) +
  geom_boxplot(aes(fill = factor(quality))) +
  ggtitle("Total Sulfur Dioxide Boxplots by Quality")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_medians_4}
total.sulfur.dioxide.by_quality <- red_wines %>%
  group_by(quality) %>%
  summarise(total.sulfur.dioxide_median = median(total.sulfur.dioxide),
            total.sulfur.dioxide_mean = mean(total.sulfur.dioxide),
            n = n())
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Conditional_means_plot_4}
ggplot(data = total.sulfur.dioxide.by_quality, aes(x = quality, y = total.sulfur.dioxide_median)) +
  geom_line(linetype = 5, colour = "yellow", size = 2) +
  ggtitle("Total Sulfur Dioxide conditional means by Quality")
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
Alcohol has a strong correlation with density; however, alcohol is the variable with the highest correlation to quality, while density has a negligible correlation with quality.
The group fixed acidity, citric acid and ph are highly correlated between them. Volatile acidity has strong correlation with only citric acid from this group; interestingly it is also the 'acidic' variable with the strongest correlation to quality.
Sulphates has a strong correlation to chlorides, but sulphates doubles its correlation to quality compared to chlorides.
Free sulfur dioxide and total sulfur dioxide also have strong correlation. Both have low correlation with quality, but total sulfur dioxide had a stronger correlation; however, looking at the conditional means plot there is no clear relationship with quality.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
Residual sugar has low correlation with all the other variables, including quality.

### What was the strongest relationship you found?
Fixed acidity and ph.


# Multivariate Plots Section

I will start exploring the relationship between volatile acidity, citric acid, and quality. 
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots_1}
ggplot(data = red_wines, aes(x = volatile.acidity, y = citric.acid)) +
  geom_point(aes(size = factor(quality), alpha = 0.1)) +
  geom_smooth(method = "lm") +
  coord_cartesian(ylim = c(0, 1)) +
  ggtitle("Volatile Acidity vs. Citric Acid, by Quality") +
  guides(alpha=guide_legend(title="alpha"))
```
On the x axis there seems to be larger dots (higher score quality) to the left; whereas on the y axis it seems that there are dots of all sizes throughout. The relationship between volatile acidity and citric acid is shown in the plot by the smooth line.

I will facet this plot.
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots_2}
ggplot(data = red_wines, aes(x = volatile.acidity, y = citric.acid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  ggtitle("Volatile Acidity vs. Citric Acid, Faceted by Quality") +
  facet_wrap(~ quality, ncol = 2)
```
The relationship between volatile acidity and citric acid is consistent for all the quality scores. Since I am trying to build a model to predict quality, I confirm my idea of using only volatile acidity in my model.

Now, volatile acidity, alcohol, and quality. Alcohol and volatile acidity are the two variables with higher correlation to quality.
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots_3}
ggplot(data = red_wines, aes(x = volatile.acidity, y = alcohol)) +
  geom_point(aes(colour = factor(quality))) +
  geom_smooth(method = "lm") +
  ggtitle("Volatile Acidity vs. Alcohol, by Quality") +
  guides(alpha=guide_legend(title="alpha"))
```

I will try a faceted plot
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots_4}
ggplot(data = red_wines, aes(x = volatile.acidity, y = alcohol)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  ggtitle("Volatile Acidity vs. Alcohol, Faceted by Quality") +
  facet_wrap(~ quality, ncol = 2)
```
The relationship between alcohol and volatile acidity, for different qualities is not consistent. Therefore I confirm the use of these two variables in my model, for I will consider them as 'independent'.

Now I will build the model:
'm1' is simply quality as a function of alcohol, 'm2' adds volatile acidity as an explanatory variable to 'm1', and 'm3' adds sulphates as an explanatory variable to 'm2'.
```{r echo=FALSE, message=FALSE, warning=FALSE, Building_the_Linear_Model}
m1 <- lm(quality ~ alcohol, data = red_wines)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + sulphates)
mtable(m1, m2, m3)
```
R^2 for m3 is only 0.336; however, all variables in the model are significant and independent.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
I could confirm the high correlation between volatile acidity and citric acid. Also, that alcohol and volatile acidity can be considered as independent varibles.

### Were there any interesting or surprising interactions between features?
Interactions confirmed the features described in previous sections.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.
Given the conditional median plots from the Bivariate Analysis section, a linear model was chosen to predict quality scores. 
There were initially five varibles of interest for this plot: alcohol, volatile acidity, sulphates, citric acid, and total sulfur dioxide. Total sulfur dioxide was discarded with its conditional median plot. Also volatile acidity and citric acid were found to have a strong relationship, given the plots for these variables and their relationship with quality, only volatile acidity was left in the model. This considerations gave rise to the 'm3' model with only three explanatory variables, alcohol, volatile acidity, and sulphates.
The dataset can be thought as relatively small, so in order to avoid overfitting to this dataset, the relatively low R^2 value can be thought as sufficient. Also, bear in mind that even though given by experts, quality scores are subjective values; and most scores are either 5 or 6. Possibly, a better model could be built with a larger ddataset, and a quality score variable with larger variance.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x = quality, y = ..density..), data = red_wines) +
  geom_density(alpha = 0.2, colour = "black", fill = "blue", size = 1.5) +
  scale_x_continuous(limits = c(2, 9), breaks = seq(2, 8, 1)) +
  ggtitle("Quality Density Plot")
```

### Description One
Most quality scores are 5 or 6. There isn't much variance in the variable I am trying to explain/predict.

### Plot Two
```{r echo=FALSE, Plot_Two}
plot1 <- ggplot(data = red_wines, aes(x = factor(quality), y = alcohol)) +
  geom_boxplot() +
  ggtitle("Alcohol Boxplots by Quality")

plot2 <- ggplot(data = red_wines, aes(x = factor(quality), y = volatile.acidity)) +
  geom_boxplot() +
  ggtitle("Volatile Acidity Boxplots by Quality")

plot3 <- ggplot(data = red_wines, aes(x = factor(quality), y = sulphates)) +
  geom_boxplot() +
  ggtitle("Sulphates Boxplots by Quality")

grid.arrange(plot1, plot2, plot3, ncol = 2)
```

### Description Two
Boxplots of the three variables of the model by quality.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggpairs(red_wines, columns = c(3,11,12,13), 
        upper = list(continuous = "smooth", 
                     params = c(alpha = 0.1,colour=I("#ffdb00"))),
        lower = list(continuous = "cor"))
```

### Description Three
Variables in the model 'm3'. 

------

# Reflection

This dataset contains several chemical characteristics of red wines, and a quality varible that contains quality scores for each of the 1599 observations.  Initially, keeping in mind that the final goal was to build a model that explained quality as a function of the chemical characteristics of red wines, a univariate exploration and analysis was conducted. The first important finding was that quality did not had too much variance; most of the scores were either 5 or 6. The fact that citric acid didn't had a normal distribution aslo was important for its exclusion of the model at the end. 

In the bivariate exploration and analysis, five potential variables were recognized as variables of interest for their correlation to quality. It was also found a strong correlation between two of the variables of interest, namely volatile acidity and citric acid. Finally, conditional median plots gave insight in the linear relationship of quality and that variables of interest, except for total sulfur dioxide.

In the last section of Multivariate exploration and analysis, the previous findings were confirmed and the model was built. 